---
title: "Assignment 1"
output: html_document
date: "2025-01-07"
author: "Jamie Chang"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### GitHub Link:

[View GitHub Repository](https://github.com/jamiechang03/BIOL432_A1)

***

### R Scripts

The following R script generates the data frame and writes it to a csv file named "measurements.csv"

```{r}
source('dataGenerato.R')
```

The following R script adds a Limb Volume column to the data, and overwrites the "measurements.csv" file with the updated data.It also loads the dplyr package. 

```{r message=FALSE}
source('volumeEstimato.R')
```

***

### Load & Check Data

```{r results="hide"}
ReportData=read.csv("measurements.csv")
head(ReportData)
tail(ReportData)
dim(ReportData)
summary(ReportData)
str(ReportData)
```

***

### Sorting Data

```{r}
ReportData = ReportData %>%
  arrange(Species, Observer, Volume) # sort data by species, then by observer, then by limb volume
```

***

### Average Species Volume Table

```{r}
avg_vol = ReportData %>%
  group_by(Species) %>% # groups by species
  summarize(Mean_Volume=mean(Volume) # calculates mean for each species
)
knitr::kable(avg_vol) # puts results into a knitted table
```

***

### Number of observers for each species/observer combination

```{r message=FALSE}
observer_n = ReportData %>%
  group_by(Species, Observer) %>% # groups by species and observer
  summarize(n_Observations = n() # counts the number of observations for each observer/species combination
) 
knitr::kable(observer_n) # puts results into a knitted table
```

***

### Boxplot

```{r fig.cap="Boxplot showcasing the distribution of limb volume among species."}
library(ggplot2)
source("https://bit.ly/theme_pub")  
theme_set(theme_pub()) 
ggplot(aes(x=Species, y=Volume), data=ReportData) + geom_boxplot() + ylab("Limb Volume")
```

***

### Final Plot

```{r fig.cap="Histogram showcasing the distribution of limb volume, faceted by species.", message=FALSE}
ggplot(data=ReportData, aes(x=Volume)) + geom_histogram() + facet_wrap(~Species) + ylab("Count") + xlab("Limb Volume")
```

























